---
title: "Test speed and accuracy of R-Rerf Iteration 9"
author: "James Browne"
date: "April 24, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, error=FALSE, message=FALSE, cache = FALSE)
source('~/dropbox/gitRepos/R-RerF/versions/iter9/rfr_function.R')
#source('~/dropbox/gitRepos/R-RerF/rfr_mod.R')
library(compiler)
library(parallel)
setCompilerOptions("optimize"=3)
rfr <- cmpfun(rfr)
error_rate <- cmpfun(error_rate)
#enableJIT(0)
```
This iteration trains the forest using 2 cores.  This iteration reduces the space requirements to store the forest data structure and improved the testing speed.

**********************************************************************
#### 10 runs iris, MinParent=6, trees=100, MaxDepth=0, bagging = .20, FUN=makeA, options=ncol(X)
**********************************************************************
```{r run-iris, cache = FALSE}
#prepare data
X <- as.matrix(iris[,1:4])
Y <- as.numeric(iris[,5])
ptmtrain <- NA
ptmtest <- NA
error <- NA
for (i in 1:10){
ptm <- proc.time()
mcrun <- function(...) rfr(X,Y, trees=50, bagging=.2)
forest<-do.call(c,mclapply(seq_len(2), mcrun, mc.cores =2))
ptmtrain[i]<- (proc.time() - ptm)[3]
ptm <- proc.time()
error[i] <- error_rate(X,Y,forest)
ptmtest[i]<- (proc.time() - ptm)[3]
cat("|")
}
cat("iris training complete:\nTraining Time Summary\n")
summary(ptmtrain)
cat("iris error complete:\nError Summary:\n")
summary(error)
cat("Test Time Summary:\n")
summary(ptmtest)
```

**********************************************************************
#### 10 runs wdbc (569x30), MinParent=6, trees=100, MaxDepth=0, bagging = .2, FUN=makeA, options=ncol(X)
**********************************************************************
```{r run-wdbc, cache = FALSE}
#prepare data
X <- read.csv("~/dropbox/gitRepos/R-RerF/data/wdbc.data", header=FALSE)
Y <- as.numeric(X[,2])
X <- X[,3:ncol(X)]
ptmtrain <- NA
ptmtest <- NA
error <- NA
for (i in 1:10){
ptm <- proc.time()
mcrun <- function(...) rfr(X,Y, trees=50, bagging=.2, MaxDepth =0)
forest<-do.call(c,mclapply(seq_len(2), mcrun))
ptmtrain[i]<- (proc.time() - ptm)[3]
ptm <- proc.time()
error[i] <- error_rate(X,Y,forest)
ptmtest[i]<- (proc.time() - ptm)[3]
cat("|")
}
cat("wdbc training complete:\nTraining Time Summary:\n")
summary(ptmtrain)
cat("wdbc error complete:\nError Summary:\n")
summary(error)
cat("Test Time Summary:\n")
summary(ptmtest)
```

**********************************************************************
#### 10 runs Easy set (2000x2), MinParent=6, trees=100, MaxDepth=0, bagging = .2, FUN=makeA, options=ncol(X)
**********************************************************************
```{r run-easy, cache = FALSE}
#prepare data
X <- read.csv("~/dropbox/gitRepos/R-RerF/data/easy_gen.csv", header = TRUE)
Y <- X[,3]+1 #don't forget to start at 1 instead of 0.
X<- X[,1:2]
ptmtrain <- NA
ptmtest <- NA
error <- NA
for (i in 1:10){
ptm <- proc.time()
mcrun <- function(...) rfr(X,Y, trees=50, bagging=.2)
forest<-do.call(c,mclapply(seq_len(2), mcrun))
ptmtrain[i]<- (proc.time() - ptm)[3]
ptm <- proc.time()
error[i] <- error_rate(X,Y,forest)
ptmtest[i]<- (proc.time() - ptm)[3]
cat("|")
}
cat("Easy set training complete:\nTraining Time Summary\n")
summary(ptmtrain)
cat("Easy set error complete:\nError Summary:\n")
summary(error)
cat("Test Time Summary:\n")
summary(ptmtest)
```

**********************************************************************
#### 10 runs Trunk (1000x10), MinParent=6, trees=100, MaxDepth=0, bagging = .2, FUN=makeA, options=ncol(X)
**********************************************************************
```{r run-trunk, cache = FALSE}
#prepare data
X <- read.csv("~/dropbox/gitRepos/R-RerF/data/Trunk/Trunk_Xtrain.dat", header = FALSE)
Y <- read.csv("~/dropbox/gitRepos/R-RerF/data/Trunk/Trunk_Ytrain.dat", header = FALSE)+1
Xt <- read.csv("~/dropbox/gitRepos/R-RerF/data/Trunk/Trunk_Xtest.dat", header = FALSE)
Yt <- read.csv("~/dropbox/gitRepos/R-RerF/data/Trunk/Trunk_Ytest.dat", header = FALSE)+1
ptmtrain <- NA
ptmtest <- NA
error <- NA
for (i in 1:10){
ptm <- proc.time()
mcrun <- function(...) rfr(X,Y, trees=50, bagging=.2)
forest<-do.call(c,mclapply(seq_len(2), mcrun))
ptmtrain[i]<- (proc.time() - ptm)[3]
ptm <- proc.time()
mcerr <- function(...) error_rate()
error[i] <- error_rate(X,Y,forest)
ptmtest[i]<- (proc.time() - ptm)[3]
cat("|")
}
cat("Trunk set training complete:\nTraining Time Summary\n")
summary(ptmtrain)
cat("Trunk set error complete:\nError Summary:\n")
summary(error)
cat("Test Time Summary:\n")
summary(ptmtest)
```
