---
title: "Random/Randomer Forest Bootstrap vs Subsample"
author: "Jaewon Chung"
output:
  html_document:
    code_folding: hide
    fig_caption: yes
    fig_height: 8
    fig_width: 8
    highlight: pygments
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 2
  keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Random Forest Bootstrap vs Subsampling
Consider the usual Random Forest/RerF algorithm where samples are bagged in order to reduce variance. This experiment compares the performance of RF/RerF algorithm when samples are bagged and subsampled (no bagging). When subsampling, we consider $1 - \frac{1}{e} \approx 0.632$ number of samples.

# Running RF
Variables swept over:

* mtrys - $p^{(1/4, 1/2, 3/4, 1)}$
* replace - T or F
* in_bag_samples = 0.632 when replace = F

```{r randomForest, echo=T, eval=FALSE}
# evaluate classifiers on benchmark datasets

rm(list = ls())
options(scipen = 999)

library(randomForest)

# Parameters

# For local
#rerfPath <- "./"
#dataPath <- "./Data/processed/"
#source(paste0(rerfPath, "R/Code/Utils/GetFolds.R"))
#source(paste0(rerfPath, "R/Code/Utils/Utils.R"))

# For MARCC
rerfPath <- "~/work/jaewon/"
dataPath <- "~/work/jaewon/RandomerForest/Data/uci/processed/"
source(paste0(rerfPath, "RandomerForest/R/Utils/GetFolds.R"))
source(paste0(rerfPath, "RandomerForest/R/Utils/Utils.R"))

classifiers <- c("rf-bag", "rf-subsample")
nCl <- length(classifiers)

seed <- 20180626L

set.seed(seed = seed)

testError <- list()
OOBError <- list()
trainTime <- list()
testTime <- list()
bestIdx <- list()
params <- list()


dataSet <- "abalone"
#dataSets <- read.table(paste0(dataPath, "names.txt"))[[1]]

# Set variables
testError[[dataSet]] <- vector(mode = "list", length = nCl)
names(testError[[dataSet]]) <- classifiers
OOBError[[dataSet]] <- vector(mode = "list", length = nCl)
names(OOBError[[dataSet]]) <- classifiers
trainTime[[dataSet]] <- vector(mode = "list", length = nCl)
names(trainTime[[dataSet]]) <- classifiers
testTime[[dataSet]] <- vector(mode = "list", length = nCl)
names(testTime[[dataSet]]) <- classifiers
bestIdx[[dataSet]] <- vector(mode = "list", length = nCl)
names(bestIdx[[dataSet]]) <- classifiers
params[[dataSet]] <- vector(mode = "list", length = nCl)
names(params[[dataSet]]) <- classifiers

# Data wrangling
X <- as.matrix(read.table(paste0(dataPath, "data/", dataSet, ".csv"), header = F, sep = ",", quote = "", row.names = NULL))

p <- ncol(X) - 1L
n <- nrow(X)

Y <- as.integer(X[, p + 1L]) + 1L
X <- X[, -(p + 1L)]

# remove columns with zero variance
X <- X[, apply(X, 2, function(x) any(as.logical(diff(x))))]
# mean-center and scale by sd
X <- scale(X)

# Get folds
fold <- GetFolds(paste0(dataPath, "cv_partitions/", dataSet, "_partitions.txt"))
nFolds <- length(fold)

print(paste0("Evaluting Dataset: ", dataSet))
cat("\n")

for (m in classifiers) {
  # Parameter tuning
  if (m == "rf-bag") {
    replace <- T
  } else if (m == "rf-subsample") {
    replace <- F
  }
  
  # Control for different number of feature selection
  if (p < 5) {
    mtrys <- 1:p
  } else {
    mtrys <- ceiling(p^c(1 / 4, 1 / 2, 3 / 4, 1))
    print(paste0("Mtrys: ", mtrys))
  }
  
  if (n >= 1000) {
    nodesize <- ceiling(n * 0.002)
  } else {
    nodesize <- 1
  }
  
  params[[dataSet]][[m]] <- list(replace = replace, mtrys = mtrys, nodesize = nodesize)
  
  size <- length(mtrys)
  OOBErrors <- matrix(as.double(rep(NA, size)), ncol = nFolds, nrow = size)
  testErrors <- matrix(as.double(rep(NA, size)), ncol = nFolds, nrow = size)
  
  print(paste0("evaluating model: ", m))
  for (fold.idx in seq.int(nFolds)) {
    print(paste0("fitting fold: ", fold.idx))
    
    data <- splitData(X, Y, fold, fold.idx)
    
    for (mtrys.idx in seq.int(length(mtrys))) {
      model <- randomForest(data$X.train, data$y.train, 
                            mtry = mtrys[mtrys.idx], 
                            replace = replace, 
                            nodesize = nodesize)
      
      OOBErrors[mtrys.idx, fold.idx] <- model$err.rate[, 1][length(model$err.rate[, 1])]
      testErrors[mtrys.idx, fold.idx] <- computePredictions(model, data$X.test, data$y.test)
    }
  }
  
  OOBError[[dataSet]][[m]] <- OOBErrors
  testError[[dataSet]][[m]] <- testErrors
}

save(OOBError, testError, file = paste0(rerfPath, "RandomerForest/R/Result/2018.06.26/", dataSet, "_2018_06_26.RData"))
```

# Running RerF
Variables swept over:

* mtrys - $p^{(1/4, 1/2, 3/4, 1, 2)}$
* sparsity - 1/p, 2/p, 3/p, 4/p, 5/p
* replacement - T or F
* in_bag_samples = 0.632 when replacement = F

```{r rerf, echo=T, eval=FALSE}
# evaluate classifiers on benchmark datasets

rm(list=ls())
options(scipen = 999)

library(rerf)
library(AUC)
library(dummies)
library(R.utils)

date <- "2018.06.30/"

## For MARCC
rerfPath <- "~/work/jaewon/"
dataPath <- "~/work/jaewon/RandomerForest/Data/uci/processed/"
source(paste0(rerfPath, "RandomerForest/R/Utils/RerFEval.R"))
source(paste0(rerfPath, "RandomerForest/R/Utils/GetCatMap.R"))
source(paste0(rerfPath, "RandomerForest/R/Utils/GetFolds.R"))

## For local
# rerfPath <- "./"
# dataPath <- "./Data/uci/processed/"
# source(paste0(rerfPath, "R/Utils/RerFEval.R"))
# source(paste0(rerfPath, "R/Utils/GetCatMap.R"))
# source(paste0(rerfPath, "R/Utils/GetFolds.R"))


testError <- list()
testAUC <- list()
OOBError <- list()
OOBAUC <- list()
trainTime <- list()
OOBTime <- list()
testTime <- list()
treeStrength <- list()
treeCorr <- list()
numNodes <- list()
bestIdx <- list()
params <- list()

dataSet <- "abalone"
fold <- GetFolds(paste0(dataPath, "cv_partitions/", dataSet, "_partitions.txt"))
nFolds <- length(fold)
X <- as.matrix(read.table(paste0(dataPath, "data/", dataSet, ".csv"), header = F, sep = ",", quote = "", row.names = NULL))
catMap <- NULL
p <- ncol(X) - 1L
p.ohe <- p
Y <- as.integer(X[, p + 1L]) + 1L
X <- X[, -(p + 1L)]
# remove columns with zero variance
X <- X[, apply(X, 2, function(x) any(as.logical(diff(x))))]
# mean-center and scale by sd
X <- scale(X)
p <- ncol(X)

## Parameters
nTrees <- 500L
min.parent <- 6L
max.depth <- ceiling(log2(nrow(X) * 0.8))
bagging <- 1 / exp(1)
replacement <- T
supervised <- 0
num.cores <- 24L
seed <- 20180629L
timeout <- 500

## Classifiers
classifiers <- c("rerf-bag", "rerf-subsample")
nCl <- length(classifiers)

testError[[dataSet]] <- vector(mode = "list", length = nCl)
names(testError[[dataSet]]) <- classifiers
testAUC[[dataSet]] <- vector(mode = "list", length = nCl)
names(testAUC[[dataSet]]) <- classifiers
OOBError[[dataSet]] <- vector(mode = "list", length = nCl)
names(OOBError[[dataSet]]) <- classifiers
OOBAUC[[dataSet]] <- vector(mode = "list", length = nCl)
names(OOBAUC[[dataSet]]) <- classifiers
trainTime[[dataSet]] <- vector(mode = "list", length = nCl)
names(trainTime[[dataSet]]) <- classifiers
OOBTime[[dataSet]] <- vector(mode = "list", length = nCl)
names(OOBTime[[dataSet]]) <- classifiers
testTime[[dataSet]] <- vector(mode = "list", length = nCl)
names(testTime[[dataSet]]) <- classifiers
numNodes[[dataSet]] <- vector(mode = "list", length = nCl)
names(numNodes[[dataSet]]) <- classifiers
bestIdx[[dataSet]] <- vector(mode = "list", length = nCl)
names(bestIdx[[dataSet]]) <- classifiers
params[[dataSet]] <- vector(mode = "list", length = nCl)
names(params[[dataSet]]) <- classifiers

print(dataSet)

for (m in classifiers) {
  if (m == "rerf-bag" || m == "rerf-subsample") {
    random.matrix <- "binary"
    if (p < 5) {
      mtrys <- c(1:p, p^2)
    } else if (p >= 5 && p <= 100) {
      mtrys <- ceiling(p^c(1/4, 1/2, 3/4, 1, 2))
    } else {
      mtrys <- ceiling(p^c(1/4, 1/2, 3/4, 1, 1.5))
    }
    sparsity <- (1:min(p-1, 5))/p
    prob <- 0.5
    
    if (m == "rerf-subsample") {
      replacement <- F
    }
  }
  
  params[[dataSet]][[m]] <- list(trees = nTrees, random.matrix = random.matrix, d = mtrys, sparsity = sparsity, prob = prob, rotate = rotate,
                                 rank.transform = rank.transform, min.parent = min.parent, max.depth = max.depth, num.cores = num.cores,
                                 seed = seed, cat.map = catMap, supervised = supervised, replacement = replacement, bagging = bagging)
  
  testError[[dataSet]][[m]] <- matrix(as.double(rep(NA, nFolds*length(sparsity)*length(mtrys)*length(supervised)*max(length(prob), 1))),
                                      nrow = nFolds, ncol = length(sparsity)*length(mtrys)*length(supervised)*max(length(prob), 1))
  testAUC[[dataSet]][[m]] <- matrix(as.double(rep(NA, nFolds*length(sparsity)*length(mtrys)*length(supervised)*max(length(prob), 1))),
                                    nrow = nFolds, ncol = length(sparsity)*length(mtrys)*length(supervised)*max(length(prob), 1))
  OOBError[[dataSet]][[m]] <- matrix(as.double(rep(NA, nFolds*length(sparsity)*length(mtrys)*length(supervised)*max(length(prob), 1))),
                                     nrow = nFolds, ncol = length(sparsity)*length(mtrys)*length(supervised)*max(length(prob), 1))
  OOBAUC[[dataSet]][[m]] <- matrix(as.double(rep(NA, nFolds*length(sparsity)*length(mtrys)*length(supervised)*max(length(prob), 1))),
                                   nrow = nFolds, ncol = length(sparsity)*length(mtrys)*length(supervised)*max(length(prob), 1))
  trainTime[[dataSet]][[m]] <- matrix(as.double(rep(NA, nFolds*length(sparsity)*length(mtrys)*length(supervised)*max(length(prob), 1))),
                                      nrow = nFolds, ncol = length(sparsity)*length(mtrys)*length(supervised)*max(length(prob), 1))
  OOBTime[[dataSet]][[m]] <- matrix(as.double(rep(NA, nFolds*length(sparsity)*length(mtrys)*length(supervised)*max(length(prob), 1))),
                                    nrow = nFolds, ncol = length(sparsity)*length(mtrys)*length(supervised)*max(length(prob), 1))
  testTime[[dataSet]][[m]] <- matrix(as.double(rep(NA, nFolds*length(sparsity)*length(mtrys)*length(supervised)*max(length(prob), 1))),
                                     nrow = nFolds, ncol = length(sparsity)*length(mtrys)*length(supervised)*max(length(prob), 1))
  numNodes[[dataSet]][[m]] <- matrix(as.double(rep(NA, nFolds*length(sparsity)*length(mtrys)*length(supervised)*max(length(prob), 1))),
                                     nrow = nFolds, ncol = length(sparsity)*length(mtrys)*length(supervised)*max(length(prob), 1))
  bestIdx[[dataSet]][[m]] <- as.integer(rep(NA, nFolds))
  
  print(params[[dataSet]][[m]])
  
  # loop over folds
  for (k in seq.int(nFolds)) {
    print(paste0("fold ", k))
    
    trainIdx <- unlist(fold[-k])
    testIdx <- fold[[k]]
    
    # evaluate models
    res <- RerFEval(X[trainIdx, ], Y[trainIdx], X[testIdx, ], Y[testIdx], params[[dataSet]][[m]], timeout = timeout)
    
    testError[[dataSet]][[m]][k, ] <- res$testError
    testAUC[[dataSet]][[m]][k, ] <- res$testAUC
    OOBError[[dataSet]][[m]][k, ] <- res$oobError
    OOBAUC[[dataSet]][[m]][k, ] <- res$oobAUC
    trainTime[[dataSet]][[m]][k, ] <- res$trainTime
    OOBTime[[dataSet]][[m]][k, ] <- res$oobTime
    testTime[[dataSet]][[m]][k, ] <- res$testTime
    numNodes[[dataSet]][[m]][k, ] <- res$numNodes
    bestIdx[[dataSet]][[m]][k] <- res$best.idx
    
    save(testError, testAUC, OOBError, OOBAUC, trainTime, OOBTime, testTime, numNodes, bestIdx, params, 
         file = paste0(rerfPath, "RandomerForest/R/Experiments/", date, "Results/", dataSet, ".RData"))
  }
}
```

# Results

The p-values corresponds to one-tailed Wilcox Ranked Test. Subsampling is not better than bagging.

```{r plotting, fig.keep = 'none'}
library('ggplot2')
library('grid')
library('gridExtra')
library('gtable')

plotResults <- function(df, classifiers, y.min = -1, y.max = 1) {
  categories <- df[['Category']][seq(1, nrow(df), 5)]
  error.cls.1 <- df[[classifiers[1]]]
  error.cls.2 <- df[[classifiers[2]]]
  
  error.cls.1 <- rowMeans(t(matrix(error.cls.1, nrow = 5)))
  error.cls.2 <- rowMeans(t(matrix(error.cls.2, nrow = 5)))
  
  # Compute One sided Wilcox Rank Test
  alt <- 'less'
  
  wilcox.all <- wilcox.test(error.cls.1, error.cls.2, paired = T, alternative = alt, exact = F)
  wilcox.categorical <- wilcox.test(error.cls.1[categories == 'categorical'], 
                                    error.cls.2[categories == 'categorical'],
                                    paired = T,
                                    alternative = alt, 
                                    exact = F)
  wilcox.numeric <- wilcox.test(error.cls.1[categories == 'numeric'],
                                error.cls.2[categories == 'numeric'], 
                                paired = T,
                                alternative = alt, 
                                exact = F)
  
  pvalue.all <- format(round(wilcox.all$p.value, 2), scientific = T)
  pvalue.categorical <- format(round(wilcox.categorical$p.value, 2), scientific = T)
  pvalue.numeric <- format(round(wilcox.numeric$p.value, 2), scientific = T)
  
  mean.error <- sqrt((error.cls.1 + error.cls.2) / 2)
  difference.error <- sqrt(abs(error.cls.1 - error.cls.2)) * sign(error.cls.1 - error.cls.2)
  
  df <- data.frame(mean.error, difference.error, categories)
  names(df) <- c("mean", "diff", "category")
  df$category <- factor(df$category)
  
  # Plot scatter
  fig <- ggplot(df, aes(x = mean, y = diff, color = category)) + geom_point() +
    theme(
      panel.background = element_blank(), axis.line = element_line(colour = "black")
    ) +
    labs(
      x = expression(sqrt("Mean Error")),
      y = expression(sqrt("Difference in Error"))
    ) +
    geom_hline(yintercept = 0) +
    xlim(0, 1) +
    ylim(y.min, y.max) +
    annotate("text", label = 'bold("Subsampling Better")', x = 1, y = y.max, parse = T, hjust = 'inward', vjust = 'inward') +
    annotate("text", label = 'bold("Subsampling Worse")', x = 1, y = y.min, parse = T, hjust = 'inward', vjust = 'inward') + 
    annotate("text", label = paste0("p=", pvalue.all), 
             x = 0, y = y.min * .6, vjust = 'inward', hjust = 'inward', color = "black") +
    annotate("text", label = paste0("p=", pvalue.categorical), 
             x = 0, y = y.min * .8, vjust = 'inward', hjust = 'inward', color = "#F8766D") +
    annotate("text", label = paste0("p=", pvalue.numeric), 
             x = 0, y = y.min, vjust = 'inward', hjust = 'inward', color = "#00BFC4")
  
  # Plot KDE
  kde <- ggplot(df, aes(x = diff, color = category)) +
    stat_density(geom = 'line', position = 'identity') + 
    stat_density(aes(x = diff, color = 'all'),  geom = 'line') +
    theme(panel.background = element_blank(), 
          axis.line = element_blank(),
          axis.ticks = element_blank(),
          axis.text = element_blank(), 
          axis.title = element_blank(), 
          legend.direction = "horizontal",
          legend.position = "bottom") +
    geom_hline(yintercept = 0) + 
    geom_vline(xintercept = 0) + 
    xlim(y.min, y.max) +
    coord_flip() + 
    scale_color_manual(values=c('#000000','#F8766D','#00BFC4'))
  
  # print(fig)
  return(list(fig = fig, kde = kde))
}


load('./../2018.07.02/uci_results.RData')
load('./../2018.07.04/df.rf.RData')

res <- plotResults(df, c('RerF','RerF.subsample'), -.6, .6)
fig.1 <- res$fig
kde.1 <- res$kde + 
  #scale_color_discrete("Dataset Type", labels = c('All', 'Categorical', 'Continuous')) +
  scale_color_manual(values = c('#000000','#F8766D','#00BFC4'),
                     labels = c("All",
                                "Categorical",
                                "Continuous"))

res <- plotResults(df.rf, c('rf.bag', 'rf.subsample'), -.22, .22)
fig.2 <- res$fig
kde.2 <- res$kde

# Get legend for separate plotting
g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

leg <- g_legend(kde.1) #+ guides(colour = guide_legend(override.aes = list(size = 3))))
# Combine figures
g.1 <- ggplotGrob(fig.1 + theme(legend.position = 'none'))
panel_id <- g.1$layout[g.1$layout$name == "panel",c("t","l")]
g.1 <- gtable_add_cols(g.1, unit(4,"cm"))
g.1 <- gtable_add_grob(g.1, ggplotGrob(kde.1 + theme(legend.position = 'none', plot.margin = unit(c(.13, 0, 0, 0), 'cm'))),
                     t = panel_id$t, l = ncol(g.1))

g.2 <- ggplotGrob(fig.2 + theme(legend.position = 'none'))
panel_id <- g.2$layout[g.2$layout$name == "panel", c("t","l")]
g.2 <- gtable_add_cols(g.2, unit(4,"cm"))
g.2 <- gtable_add_grob(g.2, ggplotGrob(kde.2 + theme(legend.position = 'none', plot.margin = unit(c(.13, 0, 0, 0), 'cm'))),
                     t = panel_id$t, l = ncol(g.2))

top <- grid.arrange(g.1, nrow = 1, top = textGrob("RerF Bootstrap - Subsampling", gp=gpar(fontface = "bold")))
bottom <- grid.arrange(g.2, nrow = 1, top = textGrob("RF Bootstrap - Subsampling", gp=gpar(fontface = "bold")))
```

```{r ggplot-2, fig.width = 8, fig.height = 8}
output <- grid.arrange(top, bottom, leg, nrow = 3, heights=c(1, 1, .1))

ggsave(filename = './result.pdf', plot = output, width = 7, height = 7)
```

